NLTK: The Natural Language Toolkit

### 2002-03-28 technical/acl2002/submission.tex
- Spellcheck - Put mbox's around urls to prevent linewrap
svn/trunk@531

% To do:
%   - Write up evaluation
%   - Write up other approaches
%   - Write up conclusion
%   - General clean-up
%   - ispell
%   - Should URLs include http:// or not (we should be consistant)?
% We assume that the user is familiar with the acronym ``NLP;'' is
% that ok?

@@

% [EL] ``functionality'' isn't a great word, here.
Teachers of introductory courses on computational linguistics are
often faced with the challenge of setting up a practical programming
component for student assignments and projects.  This is a difficult
task because different comptational linguistics domains require a
variety of different data structures and functionality; and because a
diverse range of topics may need to be included in the syllabus.

@@

% [EL] This was too ambiguous for me (e.g., we discuss Python)
%All materials discussed here are available under an open source
%license from \url{nltk.sf.net}.  NLTK runs on most platforms,
%including Windows, OS X, Linux, and UNIX.
The Natural Language Toolkit is available under an open source license
from \url{http://nltk.sf.net}.  NLTK runs on most platforms, including
Windows, OS X, Linux, and UNIX.

@@

% [EL] is ``defines'' the right word?
The toolkit is implemented as a collection of independant
\emph{modules}, each of which defines a specific data structure or
task.

A set of core modules defines basic data types and processing systems
that are used througout the toolkit.  The \texttt{token} module
provides basic classes for processing individual elements of text,
such as words or sentences.  The \texttt{tree} module defines data
structures for representing tree structures over text, such as syntax
trees and morphological trees.  The \texttt{probability} module
implements classes that encode frequency distributions and probability
distributions, including a variety of statistical smoothing
techniques.

@@

The \texttt{parser} module defines a high-level interface for
producing trees that represent the structures of texts.  The
\texttt{chunkparser} module defines a subinterface for parsers that
identify non-overlapping linguistic groups (such as base noun phrases)
in unrestricted text.

@@

The \texttt{classifier.featureselection} module defines a standard
interface for choosing which features are relevant for a particular
classification task.  Good feature selection can signifigantly improve
classification performance.

@@

% [EL] Provide some indication of how much tutorial material is
% available?  (Currently, 8 tutorials, mean=16pp, range=6pp-52pp.)
\paragraph{\textit{Tutorials}} teach students how to use the toolkit,
in the context of performing specific tasks.  Each tutorial focuses on
a single domain, such as tagging, probablistic systems, or text
classification.  The tutorials include a high-level discussion that
explains and motivates the domain, followed by a detailed
walk-through that uses examples to show how NLTK can be used to
perform specific tasks.

@@

\subsection{Assigments and Projects}

NLTK can be used to create student assignments of varying difficulty
and scope.  
% Use a module
In the simplest assignments, students experiment with an existing
module.  The wide variety of existing modules provide many opportunies
for creating these simple assignments.
% Edit/extend a module.
Once students become more familiar with the toolkit, they can be asked
to make minor changes or extensions to an existing module.
% Develop a new module.
A more challenging task is to develop a new module.  Here, NLTK
provides some useful starting points: predefined interfaces and data
structures, and existing modules that implement the same interface.
% Advanced projects..
Advanced projects typically involve the development of entirely new
functionality for a previously unsupported NLP task, or the
development of a complete system out of existing and new modules.

@@

The chunking tutorial motivates chunk parsing, describes each rule
type, and provides all the necessary code for the assignment.  The
provided code is responsible for loading the chunked, part-of-speech
tagged text using an existing tokenizer, creating an unchunked version
of the text, applying the chunk rules to the unchunked text, and
scoring the result.  Students focus on the NLP task only -- providing
a ruleset with the best coverage.

@@

The next example illustrates a brute-force statistical approach.  The
student calculated how often each part-of-speech tag was included in a
noun phrase.  They then constructed chunks from any sequence of tags
that occured in a noun phrase more than 50\% of the time.

@@

% Do we want to mention that there were 40h of lecture time?
% Should we be more explicit about who took the class?
We used NLTK as a basis for the assignments and student projects in
CIS-530, an introductory computational linguistics class taught at the
University of Pennsylvania.  CIS-530 is a graduate level class,
although some advanced undergraduates were also enrolled.  Most
students had a background in either computer science or linguistics
(but not both).  Students were required to complete five assignments,
two exams, and a final project.  All class materials are available
from the corse website \url{http://www.cis.upenn.edu/~cis530/}.

@@

We did encounter a few difficulties during the semester.  One problem
was finding large clean corpera that the students could use for their
assignments.  Several of the students needed assistance finding
suitable corpera for their final projects.  Another issue was the fact
that we were actively developing NLTK during the semester; some
modules were only completed one or two weeks before the students used
them.  As a result, students who worked at home needed to download new
versions of the toolkit several times throughout the semester.
Luckily, Python has extensive support for installation scripts, which
made these upgrades much simpler.  The students encountered a couple
of bugs in the toolkit, but none were serious, and all were quickly
corrected.

@@

%Future work: 
We plan to continue extending the breadth of materials covered by the
toolkit.  We are currently working on NLTK modules for Hidden Markov
Models, probablistic parsing, and tree adjoining grammars.  We also
plan to increase the number of algorithms implemented by some existing
modules, such as the text classification module.

Finding suitable corpera is a prerequisite for many studnet
assignments and projects.  We are therefore putting together a
standard collection of large, clean, open-source corpera, containing
corpera appropriate for every module defined by the toolkit.

>>>>

% To do:
%   - Write up other approaches
%   - General clean-up
% We assume that the user is familiar with the acronym ``NLP;'' is
% that ok?

@@

% [EL] ``functionality'' isn't a great word, here.
Teachers of introductory courses on computational linguistics are
often faced with the challenge of setting up a practical programming
component for student assignments and projects.  This is a difficult
task because different computational linguistics domains require a
variety of different data structures and functionality; and because a
diverse range of topics may need to be included in the syllabus.

@@

% [EL] This was too ambiguous for me (e.g., we discuss Python)
%All materials discussed here are available under an open source
%license from \url{nltk.sf.net}.  NLTK runs on most platforms,
%including Windows, OS X, Linux, and UNIX.
The Natural Language Toolkit is available under an open source license
from \mbox{\url{http://nltk.sf.net/}}.  NLTK runs on most platforms, including
Windows, OS X, Linux, and UNIX.

@@

% [EL] is ``defines'' the right word?
The toolkit is implemented as a collection of independent
\emph{modules}, each of which defines a specific data structure or
task.

A set of core modules defines basic data types and processing systems
that are used throughout the toolkit.  The \texttt{token} module
provides basic classes for processing individual elements of text,
such as words or sentences.  The \texttt{tree} module defines data
structures for representing tree structures over text, such as syntax
trees and morphological trees.  The \texttt{probability} module
implements classes that encode frequency distributions and probability
distributions, including a variety of statistical smoothing
techniques.

@@

The \texttt{parser} module defines a high-level interface for
producing trees that represent the structures of texts.  The
\texttt{chunkparser} module defines a sub-interface for parsers that
identify non-overlapping linguistic groups (such as base noun phrases)
in unrestricted text.

@@

The \texttt{classifier.featureselection} module defines a standard
interface for choosing which features are relevant for a particular
classification task.  Good feature selection can significantly improve
classification performance.

@@

% [EL] Provide some indication of how much tutorial material is
% available?  (Currently, 8 tutorials, mean=16pp, range=6pp-52pp.)
\paragraph{\textit{Tutorials}} teach students how to use the toolkit,
in the context of performing specific tasks.  Each tutorial focuses on
a single domain, such as tagging, probabilistic systems, or text
classification.  The tutorials include a high-level discussion that
explains and motivates the domain, followed by a detailed
walk-through that uses examples to show how NLTK can be used to
perform specific tasks.

@@

\subsection{Assignments and Projects}

NLTK can be used to create student assignments of varying difficulty
and scope.  
% Use a module
In the simplest assignments, students experiment with an existing
module.  The wide variety of existing modules provide many opportunities
for creating these simple assignments.
% Edit/extend a module.
Once students become more familiar with the toolkit, they can be asked
to make minor changes or extensions to an existing module.
% Develop a new module.
A more challenging task is to develop a new module.  Here, NLTK
provides some useful starting points: predefined interfaces and data
structures, and existing modules that implement the same interface.
% Advanced projects..
Advanced projects typically involve the development of entirely new
functionality for a previously unsupported NLP task, or the
development of a complete system out of existing and new modules.

@@

The chunking tutorial motivates chunk parsing, describes each rule
type, and provides all the necessary code for the assignment.  The
provided code is responsible for loading the chunked, part-of-speech
tagged text using an existing tokenizer, creating an unchunked version
of the text, applying the chunk rules to the unchunked text, and
scoring the result.  Students focus on the NLP task only -- providing
a rule set with the best coverage.

@@

The next example illustrates a brute-force statistical approach.  The
student calculated how often each part-of-speech tag was included in a
noun phrase.  They then constructed chunks from any sequence of tags
that occurred in a noun phrase more than 50\% of the time.

@@

% Do we want to mention that there were 40h of lecture time?
% Should we be more explicit about who took the class?
We used NLTK as a basis for the assignments and student projects in
CIS-530, an introductory computational linguistics class taught at the
University of Pennsylvania.  CIS-530 is a graduate level class,
although some advanced undergraduates were also enrolled.  Most
students had a background in either computer science or linguistics
(but not both).  Students were required to complete five assignments,
two exams, and a final project.  All class materials are available
from the course website \mbox{\url{http://www.cis.upenn.edu/~cis530/}}.

@@

We did encounter a few difficulties during the semester.  One problem
was finding large clean corpora that the students could use for their
assignments.  Several of the students needed assistance finding
suitable corpora for their final projects.  Another issue was the fact
that we were actively developing NLTK during the semester; some
modules were only completed one or two weeks before the students used
them.  As a result, students who worked at home needed to download new
versions of the toolkit several times throughout the semester.
Luckily, Python has extensive support for installation scripts, which
made these upgrades much simpler.  The students encountered a couple
of bugs in the toolkit, but none were serious, and all were quickly
corrected.

@@

%Future work: 
We plan to continue extending the breadth of materials covered by the
toolkit.  We are currently working on NLTK modules for Hidden Markov
Models, probabilistic parsing, and tree adjoining grammars.  We also
plan to increase the number of algorithms implemented by some existing
modules, such as the text classification module.

Finding suitable corpora is a prerequisite for many student
assignments and projects.  We are therefore putting together a
standard collection of large, clean, open-source corpora, containing
corpora appropriate for every module defined by the toolkit.
